My title

### 2024-03-11 paper/paper.qmd
Made the model

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
# 
# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
# Load the required library
library(MASS)
library(ggplot2)
library(dplyr)
library(cowplot)

# Check the column names in the merged data
colnames(merged_data)

# Make sure the column name for Total Deaths matches
# Adjust the column name accordingly if needed

# Create a negative binomial model
model_nb <- glm.nb(`Total.Deaths` ~ `OriginalValue` + `Cause`, data = merged_data)

# View the summary of the negative binomial model
summary(model_nb)

# Load the required library
library(ggplot2)

# Predict the values using the negative binomial model
merged_data$predicted_deaths <- predict(model_nb, type = "response")

# Plot the relationship between Air Quality and Predicted Deaths
ggplot(merged_data, aes(x = OriginalValue, y = predicted_deaths, color = Cause)) +
  geom_point() +
  geom_smooth(method = "glm.nb", formula = y ~ x, se = FALSE) +
  labs(x = "Air Quality Health Index", y = "Predicted Deaths", color = "Cause") +
  theme_minimal()

# Assuming you have loaded your data into 'air_quality_data' and 'deaths_data'

# Merge the air quality and deaths data by year
merged_data <- merge(air_quality_data, deaths_data, by = "Year")

# Plot the regression graph
regression_plot <- ggplot(merged_data, aes(x = OriginalValue, y = Total.Deaths)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line
  labs(title = "Regression Graph: Air Quality vs Deaths",
       x = "Air Quality Index",
       y = "Total Deaths") +
  theme_minimal()

# Show the plot
print(regression_plot)

library(ggplot2)
library(tidyr)

# Your dataset
data <- data.frame(
  "Year" = c(2000:2021),
  "Category" = c(17.6, 17.5, 17.7, 19.9, 17.6, 12.9, 14.9, 14, 14.3, 16.1, 28, 26.9, 21.3, 18.1, 25.7, 31.9, 31.2, 25.8, 52, 21.7, 15.8, 32.9),
  "Provincial Average" = c(13.5, 9.5, 11.5, 11.5, 13.6, 8.9, 11.7, 10.3, 9.8, 11.9, 16.3, 11.3, 13.3, 10.3, 16.7, 17, 11.9, 14.9, 32.8, 15.3, 10.8, 26.2),
  "10th Percentile" = c(23.7, 22.4, 24.5, 33.5, 21.6, 16.6, 18.4, 16.8, 19.9, 22.4, 41.4, 60.5, 31.5, 24, 34.1, 48.1, 49.4, 34.9, 65.9, 28, 20.1, 42.4),
  "Calgary" = c(22, 20.8, 18.8, 35.5, 17.6, 13.2, 18.3, 16.2, 14.5, 19.5, 30.8, 24, 21.9, 20.8, 21.2, 28.1, 14.7, 34.6, 58.1, 19.2, 18.9, 44.2),
  "Edmonton" = c(24.9, 24, 23.4, 24, 21.4, 14.4, 17.8, 14.8, 18.6, 21.5, 46.4, 26.9, 22.3, 29.5, 26, 23.7, 22.7, 29, 49.1, 26.2, 20.6, 38.4),
  "Fort McMurray" = c(13.7, 15.9, 15.2, 11.5, 1.5, 12.7, 14.1, 15.1, 14.9, 13.9, 16.4, 72.4, 20.8, 17, 35.1, 40.9, 208.4, 15.9, 35.4, 21.5, 12.6, 25.9),
  "Grande Prairie" = c(NA, NA, NA, NA, 14.2, 14.8, 16.4, 11.9, 11.5, 14.7, 29.1, 20.3, 19.5, 17.6, 35.9, 20.9, 21, 26.3, 60.5, 23.8, 17.9, 29.7),
  "Lethbridge" = c(NA, NA, NA, NA, 15.7, 10.1, 12.3, 12, 9.8, NA, 25.7, 18.5, 23.4, NA, 16.5, 42, 10.7, 44.9, 58.4, 20.8, 20.5, 38.2),
  "Medicine Hat" = c(NA, NA, NA, NA, 13.8, 8.8, 9.5, 10.5, 10, 24.3, NA, 18, 22.3, 34.5, 20.4, 24.5, 16.5, 29.3, 43, 13.7, 16.3, 32.8),
  "Red Deer" = c(27, 20.9, 15.3, 20.9, 27, 27, 22.3, 11.8, 11.6, 27, 37.5, 34.1, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27),
  "CAAQS" = c(27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27, 27)
)

# Reshape the data from wide to long format for plotting
data_long <- pivot_longer(data, cols = -Year, names_to = "Variable", values_to = "Value")

# Plot the average air index and total death rates over the years
graph <- ggplot(data_long, aes(x = Year, y = Value, color = Variable)) +
  geom_line() +
  labs(title = "Average Air Index and Total Death Rates Over the Years",
       x = "Year",
       y = "Value") +
  theme_minimal() +
  scale_color_manual(values = c("red", "blue", "green", "orange", "purple", "black", "brown", "lightblue","pink", "darkgreen", "lightgreen")) # Adjust color palette as needed

# Show the plot
print(graph)
```

@@

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

>>>>

# analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
# 
# analysis_data |> 
#   ggplot(aes(x = width, y = length)) +
#   geom_point(alpha = 0.8) +
#   theme_minimal() +
#   labs(x = "Wing width (mm)",
#        y = "Wing length (mm)")
```

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))

poisson_model <-
  readRDS(file = here::here("models/poisson_model.rds"))

causes_model <-
  readRDS(file = here::here("models/causes_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  #statistic = "mad",
  fmt = 2
)

modelsummary::modelsummary(
  list(
    "First model" = poisson_model
  ),
  #statistic = "mad",
  fmt = 2
)

modelsummary::modelsummary(
  list(
    "First model" = causes_model
  ),
  #statistic = "mad",
  fmt = 2
)
```

@@

# Create the posterior predictive check plot
pp_plot <- pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

# Create the posterior vs prior plot
posterior_plot <- posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```
